---
title: "Dimensionality Reduction"
author: "Khang Thai, David Park, Jonathan Ho, David Favela"
output: html_document

---


## Reading in the Dataset
```{r}
library(caret)
Invistico_Airline <- read.csv("Invistico_Airline.csv", header=TRUE)
str(Invistico_Airline)
```
## Convert satisfaction into a factor
```{r}
Invistico_Airline$satisfaction <- as.factor(Invistico_Airline$satisfaction)
```

## Create New Columns Rating Sum & Rating Mean
```{r}
Invistico_Airline$ratingSum <- as.numeric(apply(Invistico_Airline[,8:21], 1, sum))
Invistico_Airline$ratingMean <- c(Invistico_Airline$ratingSum/14)
```

## Create train & test sets
```{r}
i <- sample(1:nrow(Invistico_Airline), 0.8*nrow(Invistico_Airline), replace = FALSE)
trainAirline <- Invistico_Airline[i,]
testAirline <- Invistico_Airline[-i,]
set.seed(3)
```

## Clean out columns not needed
```{r}
trainAirline <- trainAirline[,c(4,7,25,1)]
testAirline <- testAirline[,c(4,7,25,1)]
```

## PCA
```{r}
pca_out <- preProcess(trainAirline[,1:3], method=c("center", "scale", "pca"))
pca_out
```
## Plotting the PCA with the Test Data
```{r}
train_pc <- predict(pca_out, trainAirline[,1:3])
test_pc <- predict(pca_out, testAirline[,])
plot(test_pc$PC1, test_pc$PC2, pch=c(23,21,22)[unclass(test_pc$satisfaction)], bg=c("red", "green", "blue")[unclass(testAirline$satisfaction)])
```
## Finding the accuracy of PCA
```{r}
train_df <- data.frame(train_pc$PC1, train_pc$PC2,  trainAirline$satisfaction)
test_df <- data.frame(test_pc$PC1, test_pc$PC2, testAirline$satisfaction)
library(class)

pred <- knn(train = train_df[,1:2], test = test_df[,1:2], cl=train_df[,3], k=3)
mean(pred==testAirline$satisfaction)
```
## Regression comparison to PCA
```{r}
glm1 <- glm(satisfaction~., data=trainAirline, family=binomial)
```

```{r}
probs <- predict(glm1, newdata = testAirline, type="response")
pred <- ifelse(probs>0.5,2,1)
acc1 <- mean(pred==as.integer(testAirline$satisfaction))
print(paste("glm1 accuracy = ", acc1))


```
## Results
The amount of accuracy lost from Regression to PCA was about 5% accuracy. Regression had a 73% accuracy while PCA had a 68% accuracy.

## LDA
```{r}
library(MASS)
lda1 <- lda(satisfaction~., data=trainAirline)
lda1$means
```
## Predicting the satisfaction based on the test data
```{r}
lda_pred <- predict(lda1, newdata=testAirline, type="class")
head(lda_pred$class)
```
## Regression Comparison to LDA
```{r}
glm1 <- glm(satisfaction~., data=trainAirline, family=binomial)
```

```{r}
probs <- predict(glm1, newdata = testAirline, type="response")
pred <- ifelse(probs>0.5,2,1)
acc1 <- mean(pred==as.integer(testAirline$satisfaction))
print(paste("glm1 accuracy = ", acc1))
table(pred, as.integer(testAirline$satisfaction))
```

```{r}
mean(lda_pred$class==testAirline$satisfaction)
```

## Results
Here, the accuracy for LDA was 73.44% while the accuracy of regression was 73.49% so the amount of accuracy lost was only .05%.


